### 2023-06-09 00_dataPrepReport.Rmd
clarified UTM zone

---
title: "Data preparation report"
author: "Kaija Gahm"
date: "2023-05-22"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE, message = FALSE, cache = FALSE)
```

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
```

### Step 1: Download raw data from Movebank

```{r download-movebank-data}
# base::load("movebankCredentials/pw.Rda")
# MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
# rm(pw)
minDate <- "2020-09-01 00:00"
maxDate <- "2022-12-15 11:59"
#dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
#write_feather(dat, "data/dat.feather")
dat <- read_feather("data/dat.feather")
```

Downloaded data from Movebank from `r minDate` to `r maxDate`. The
initial data has **`r nrow(dat)`** rows and `r ncol(dat)` columns. There
are **`r length(unique(dat$trackId))`** unique individual ID's in the
data at this point.

### Step 2: Fix vulture IDs

```{r fix-trackId-Nili_id}
## fix trackId. Marta told me that E03w was sometimes coded as E03, so I fixed that right up front. 
dat2 <- dat %>%
  mutate(trackId = as.character(trackId),
         trackId = case_when(trackId == "E03" ~ "E03w",
                             TRUE ~ trackId))
# number of unique individuals
length(unique(dat2$trackId)) # 126--makes sense that we lost one because one individual had a typo.

# Read in the who's who table and prepare to join the Nili_ids
ww <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, Movebank_id) %>%
  distinct()

# Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
ww %>% filter(Movebank_id %in% dat2$trackId) %>% 
  dplyr::select(Nili_id, Movebank_id) %>% 
  distinct() %>% group_by(Nili_id) %>% 
  summarize(n = length(unique(Movebank_id))) %>% 
  arrange(desc(n))
ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili_id's

all(dat2$trackId %in% ww$Movebank_id) # true

dat2 <- left_join(dat2, ww, by = c("trackId" = "Movebank_id"))
dim(dat2) # good, same number of rows.
length(unique(dat2$Nili_id))
```

Next, we join in the Who's Who table in order to fix the Movebank
trackId's and associate them with individual vultures (as defined by the
"Nili_id" column). After joining in the Nili_ids, we now have a data
frame with **`r nrow(dat2)`** rows and `r ncol(dat2)` columns. There are
now **`r length(unique(dat2$Nili_id))`** unique individual vultures
identified--the change from **`r length(unique(dat$trackId))`** is
because there were a couple cases of different trackId's being assigned
to the same Nili_id, such as "Y26" and "Y26b". No individuals are
actually lost here.

### Step 3: Remove invalid periods (cages etc.)

Next, we remove periods of time when specific vultures were in cages or
in the hospital, or when tags were not attached to the bird, etc. This
is based on the "periods_to_remove" sheet in the who's who file.

```{r remove-periods}
## annotate the data with periods to remove
toRemove <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")

toRemove <- toRemove %>%
  dplyr::select(Nili_id,
                remove_start,
                remove_end,
                reason) %>%
  mutate(across(c(remove_start, remove_end), .fns = function(x){
    lubridate::ymd(x)
  })) %>%
  dplyr::filter(!is.na(remove_end))

toRemove_long <- toRemove %>%
  group_by(Nili_id, reason) %>%
  # sequence of daily dates for each corresponding start, end elements
  dplyr::mutate(dateOnly = map2(remove_start, remove_end, seq, by = "1 day")) %>%
  # unnest the list column
  unnest(cols = c(dateOnly)) %>%
  # remove any duplicate rows
  distinct() %>%
  dplyr::select(-c(remove_start, remove_end)) %>%
  rename("status" = reason)

# Join to the original data
datAnnot <- dat2 %>%
  left_join(toRemove_long, by = c("Nili_id", "dateOnly")) %>%
  mutate(status = replace_na(status, "valid"))
nrow(datAnnot) == nrow(dat2) #T: good, same number of rows.

# Actually REMOVE the periods to remove...
datAnnot <- datAnnot %>%
  filter(status == "valid")
```

```{r remove-cages}
# Remove cages using Marta's data
# Identify and remove capture dates using Marta's code --------------------
## To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function). Then, if the bird roosted within 500 m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset. This has been validated. Note: this protocol doesn't work for the Carmel because of the position of the roost sites relative to cages. But we're just dealing with southern individuals, so that's okay.
roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
save(roosts, file = "data/roosts.Rda")
load("data/roosts.Rda")

# Identify the period of time during which the capture sites are open (when we need to do this exclusion)
start.day <- 01
start.month <- 08
end.day <-  30
end.month <- 11
distance <- 500 # distance, in meters, to calculate from the cage

# Load the information about the capture sites
sites <- read.csv("data/capture_sites.csv")

# Subset the roost dataset with the start and end dates for the capture period
sub.roosts <- roosts %>%
  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
                              format="%d-%m-%Y"),
         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
                            format="%d-%m-%Y")) %>%
  filter(date >= start_date & date <= end_date)

unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.

# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat), 
               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates

DistanceMat <- matrix(ncol = nrow(sites), nrow = nrow(crds)) 
colnames(DistanceMat) <- unique(sites$name)

for(i in 1:nrow(crds)){ # for each roost point...
  DistanceMat[i,] <- round(geosphere::distm(crds[i,], sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
  #print(i)
}

ClosestCaptureSite <- colnames(DistanceMat)[apply(DistanceMat,1,which.min)] # ID of closest capture site
ClosestCaptureDist <- apply(DistanceMat,1,min) # Distance from closest capture site

sub.roosts <- cbind(sub.roosts, ClosestCaptureSite, ClosestCaptureDist) 
sub.roosts$Captured <- ifelse(sub.roosts$ClosestCaptureDist <= distance, "Yes", "No")

sub.captured <- subset(sub.roosts, Captured == "Yes") # get list of dates which the bird was inside the cage

sub.captured.dates <- subset(sub.captured, 
                             select = c("Nili_id",
                                        "date", 
                                        "ClosestCaptureSite",
                                        "ClosestCaptureDist", 
                                        "Captured"))

## This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it. So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data

sub.captured.no.carmel <- subset(sub.captured.dates, ClosestCaptureSite != "Carmel")
sub.captured.carmel <- subset(sub.captured.dates, ClosestCaptureSite == "Carmel")

AllCarmelDates <- read.csv("data/all_captures_carmel_2010-2021.csv")
AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")

AllCarmelDates.1 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-1)))
AllCarmelDates.2 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-2)))
AllCarmelDates.3 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-3)))

AllCarmelDates.all <- rbind(AllCarmelDates, AllCarmelDates.1, AllCarmelDates.2, AllCarmelDates.3)

sub.captured.carmel <- sub.captured.carmel %>%
  mutate(known_capture = ifelse(date %in% AllCarmelDates.all$Date, 1, 0),
         Captured = ifelse(known_capture == 1 & ClosestCaptureDist <= 50, "yes", "no")) %>%
  filter(Captured == "yes") %>%
  dplyr::select(-c(known_capture))

names(sub.captured.no.carmel)
names(sub.captured.carmel)

sub.captured.dates <- rbind(sub.captured.no.carmel, sub.captured.carmel)

# We also need to exclude the day after the bird was captured
sub.captured.dates.1 <- sub.captured.dates
sub.captured.dates.1$date <- sub.captured.dates.1$date+1

sub.captured.dates <- rbind(sub.captured.dates, sub.captured.dates.1)
sub.captured.dates <- sub.captured.dates %>% 
  dplyr::distinct(Nili_id, date, .keep_all = T)

# It all looks ok, so we can subset the dataset to exclude the capture dates
datAnnot_noCaptures <- datAnnot %>%
  left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
nrow(datAnnot) == nrow(datAnnot_noCaptures) # should be TRUE. NOW we can filter.
datAnnot_noCaptures <- datAnnot_noCaptures %>%
  dplyr::filter(Captured != "Yes"|is.na(Captured)) # remove the individual*days when they were captured

datAnnot_noCaptures <- datAnnot_noCaptures %>%
  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))

# How many rows and individuals did we remove?
before <- nrow(datAnnot)
after <- nrow(datAnnot_noCaptures)
(propChange <- (before-after)/before) # approx. 1% of data removed.
length(unique(datAnnot$Nili_id))
length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.
```

The resulting data has dropped from **`r nrow(dat2)`** to
**`r nrow(datAnnot_noCaptures)`** rows, a
**`r round((nrow(dat2)-nrow(datAnnot_noCaptures))*100/nrow(dat2), 3)`% decrease**.
It has `r ncol(dat2)` columns and
**`r length(unique(datAnnot_noCaptures$Nili_id))`** unique individuals.

### Step 4: Attach birth year and sex info from who's who

```{r join-age-sex}
# Attach age and sex information
as <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, birth_year, sex) %>%
  distinct()

datAnnot2 <- datAnnot_noCaptures %>%
  dplyr::select(-c("sex")) %>%
  left_join(as, by = "Nili_id")
nrow(datAnnot2) == nrow(datAnnot_noCaptures)
length(unique(datAnnot2$Nili_id)) # 125
```

No change in the number of rows here: `r nrow(datAnnot2)` rows and
`r ncol(datAnnot2)` columns.

```{r}
datAnnot <- datAnnot2 # a hack so the rest of the code will work
```

### Step 5: Clean the data using the cleanData function.

This function includes several data cleaning steps: - Remove as invalid
data points that have zeroes in `ground_speed`, `barometric_height`, and
`external_temperature`. - Remove points that took to long to get a GPS
fix (\>89) - Remove as invalid points with heading data \<0 or \>360
degrees - Remove points with fewer than 3 satellites (because 3 are
needed for triangulation) - Remove points that "jump" unrealistically
across the map, as calculated by distance/time differences between
consecutive points - Set to NA any altitude values that jump
unrealistically between consecutive points - Remove points that "moved"
farther than expected during the night - Remove **individuals** where
less than 1/3 of their duration tracked fell inside the rough
Israel-region geographic mak - For any individuals remaining, remove
**points** that fell outside of the geographic mask (e.g. to remove
long-range forays).

```{r clean-data, echo = T, include = T}
# Clean the data
## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
mask <- sf::st_read("data/CutOffRegion.kml")
datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) %>% st_drop_geometry()
save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
load("data/datAnnotCleaned.Rda")
```

After cleaning, the new data has **`r nrow(datAnnotCleaned)`** rows (a
**`r round((nrow(datAnnot)-nrow(datAnnotCleaned))*100/nrow(datAnnot), 3)`%
decrease**) and `r ncol(datAnnotCleaned)` columns.
**`r length(unique(datAnnot$Nili_id))-length(unique(datAnnotCleaned$Nili_id))`**
individuals were removed, for a remaining total of
**`r length(unique(datAnnotCleaned$Nili_id))`** individuals.

### Step 6: Fix times and split into seasons

First, I changed the time zone to Israel time instead of UTC, since this
will make a difference to where the date line falls. Then I
re-calculated the `dateOnly` column based on the new date/times.

Next, I split the data into fall/breeding/summer seasons based on
Marta's cutoff dates: September 15th, December 15th, and May 15th. I
fixed the vultures' ages to adjust for the breeding season and then
classified them as adults or sub-adults/juveniles. Finally, I split the
data into a list, with one item per season.

```{r fix-dates}
# Fix time zone so dates make sense ---------------------------------------
## Overwrite the dateOnly column from the new times
datAnnotCleaned <- datAnnotCleaned %>%
  mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
         dateOnly = lubridate::date(timestampIsrael),
         month = lubridate::month(dateOnly),
         year = lubridate::year(dateOnly))
```

```{r split-seasons}
# Split into Marta's 3 seasons --------------------------------------------
s1 <- datAnnotCleaned %>%
  mutate(start_breeding = ifelse(month %in% c(1:5),
                                 paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
                                 paste(as.character(year), "-", "12", "-", "15", sep = "")),
         start_summer = ifelse(month == 12,
                               paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
                               paste(as.character(year), "-", "05", "-", "15", sep = "")),
         start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
         start_breeding = as.Date(start_breeding),
         start_summer = as.Date(start_summer),
         season = case_when(
           dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
           dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
           dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
  filter(dateOnly != "2022-12-15") # remove December 15 2022, because it turns out that's a non-inclusive boundary (oops)

s2 <- s1 %>%
  mutate(age = year - birth_year,
         sex = case_when(sex == "f/m" ~ "u",
                         TRUE ~ sex)) %>%
  mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
  group_by(Nili_id, year, month) %>%
  mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
  mutate(age_group = case_when(age == 0 ~ "juv",
                               age >= 1 & age <= 4 ~ "sub",
                               age >= 5 ~ "adult",
                               TRUE ~ NA)) %>%
  mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
                                  season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
                                  season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
  # explicitly set the factor levels in temporal order
  mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall")),
         season = factor(season, levels = c("breeding", "summer", "fall")))
table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read

# remove columns that were used in the season calculation
s3 <- s2 %>%
  dplyr::select(-c(month, start_breeding, start_summer, start_fall))

# Separate the seasons -----------------------------
seasons <- s3 %>%
  group_by(seasonUnique) %>%
  group_split(.keep = T)
# extract the season names in case we need a separate vector at some point.
seasonNames <- map_chr(seasons, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
```

```{r save-soc-data}
# Save copies for social analysis
# The next step will be to remove individuals with too few points per day or too few days tracked. But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric. So, before I filter for ppd and for days tracked, going to save a copy to use for social analysis. 
seasons_forSoc <- seasons
#save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
```

There are `r length(seasonNames)` seasons: `r seasonNames` (we will end
up removing Summer 2020 because the Ornitela tags started partway
through that season.) `seasons` is therefore a list of length
**`r length(seasons)`**, with the following numbers of rows in each
season: `r map_dbl(seasons, nrow)`. This data is saved as
`seasons_forSoc.Rda` so it can be used for the social network analysis
without any further changes made.

### Step 7: Get roosts

Now I use Marta's roost code, implemented in vultureUtils as
`get_roosts_df`, to get roost locations for each individual on each
night.

```{r get-roosts}
# Get roosts for each season ----------------------------------------------
# forRoosts <- map(seasons, ~.x %>% sf::st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE))
# roosts_seasons <- purrr::map(forRoosts, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id")) 
# roosts_seasons <- roosts_seasons %>%
#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
# save(roosts_seasons, file = "data/roosts_seasons.Rda")
load("data/roosts_seasons.Rda")
```

### Step 8: Remove nighttime points

Just for the data I'm going to use for movement analyses, I want to
exclude any points that occur at night. This will also help deal with
the problem that there are some individuals that seem to have way too
many points per day

```{r remove-night}
# Remove nighttime points -------------------------------------------------
before <- map_dbl(seasons, nrow)
seasons <- map(seasons, ~{
  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
                                     lat = 31.434306, lon = 34.991889,
                                     keep = c("sunrise", "sunset")) %>%
    dplyr::select("dateOnly" = date, sunrise, sunset)

  .x <- .x %>%
    dplyr::mutate(dateOnly = lubridate::ymd(dateOnly)) %>%
    dplyr::left_join(times, by = "dateOnly") %>%
    dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
    dplyr::select(-c(sunrise, sunset)) %>%
    dplyr::filter(daylight == "day")
})
after <- map_dbl(seasons, nrow)
change <- after-before
changeDF <- data.frame(seasonName = seasonNames, nBefore = before, nAfter = after) %>%
  mutate(propRowsRemoved = round((before-after)/before, 2))
```

```{r include = T, echo = F}
print(changeDF)
```

### Step 9: Include only "southern individuals" in the movement calculations

For each individual in each season, I calculated the centroid of its
coordinates. Then I looked at some histograms of the centroids in each
season and decided that 3550000 latitude (in UTM zone 36N coordinates--https://epsg.io/32636) would be a
cutoff point that would neatly divide individuals in the southern
population from the northern population (there was a large gap in the
histogram consistently around this value).

I filtered each season down to remove individuals classified as
"northern" for that season, leaving only those in the Negev population.
Keep in mind:

1)  A given individual could be "northern" in one season but "southern"
    in another season, theoretically. It would be removed only in the
    seasons when it falls in the north.

2)  Any removed individuals can still contribute to the social network
    that I will calculate--that is, they can contribute to the
    degree/strength/etc. of southern birds. I'm only removing them for
    the movement calculations, so they won't be focal individuals in the
    model.

```{r remove-northern}
# Restrict to southern individuals ----------------------------------------
# Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, https://epsg.io/32636)
## Transform to SF object, so we can get centroids
seasonsSF <- map(seasons, ~.x %>%
                   sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>%
                   sf::st_set_crs("WGS84") %>%
                   sf::st_transform(32636))

## Get centroids, so we can see who's "southern" for that season.
centroids <- map(seasonsSF, ~.x %>%
                   group_by(Nili_id) %>%
                   summarize(geometry = st_union(geometry)) %>%
                   st_centroid()) # XXX what is the actual definition of the centroid?

## Examine a histogram of centroid latitudes 
# walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here.

## Get southern individuals for each season, so we can filter the data
southernIndivs <- map(centroids, ~.x %>%
                        filter(st_coordinates(.)[,2] < 3550000) %>%
                        pull(Nili_id))

## Remove individuals not in the south
before <- seasons
seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y))
after <- seasons
beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id)))
afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id)))
df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs)
df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2)
```

After doing this filtering, the following numbers of rows and
individuals remained in each season:

```{r include = T, echo = F}
print(df)
```

### Step 10: Remove vulture\*days based on number of points per day

```{r view-ppd}
# Include only individuals with enough points per day ------------------------
# How many points per day are we looking at?
ppd <- map_dfr(seasons, ~.x %>%
             st_drop_geometry() %>%
             group_by(seasonUnique, season, Nili_id, dateOnly) %>%
             summarize(n = n()))

## Theoretically, should only have 72 points per day--12 hours, 6 points per hour. That should vary a bit. I don't really understand why so many individuals/days have more than 100 points per day, but let's proceed for now: going to downsample that to 10 minue intervals anyway, so it won't matter.
```

We had initially decided to remove vulture\*days with \< 30 points per
day. Here's what that would look like:

```{r include = T, echo = F}
# 
thresh <- 30
ppd %>%
  mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
  ggplot(aes(x = n))+
  geom_histogram(aes(fill = Status))+
  facet_wrap(~seasonUnique, scales ="free")+
  geom_vline(aes(xintercept = thresh), col = "red")+
  scale_fill_manual(values = c("darkgray", "red"))+
  theme_classic()+
  ggtitle(paste("n removed =", thresh, sep = " "))+
  ylab("# vulture*days")+
  xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
```

Restricting to days with 30 points per day ends up removing a huge
number of days, especially in the breeding season (winter) because of
lack of sunlight etc. After talking with Marta and Noa, I decided to
remove just 10 points per day. That would look like this:

```{r include = T, echo = F}
# 
thresh <- 10
ppd %>%
  mutate(Status = ifelse(n < thresh, "removed", "kept")) %>%
  ggplot(aes(x = n))+
  geom_histogram(aes(fill = Status))+
  facet_wrap(~seasonUnique, scales ="free")+
  geom_vline(aes(xintercept = thresh), col = "red")+
  scale_fill_manual(values = c("darkgray", "red"))+
  theme_classic()+
  ggtitle(paste("n removed =", thresh, sep = " "))+
  ylab("# vulture*days")+
  xlab("# points per vulture*day") # short days in breeding season, and vultures sitting on nests
```

```{r}
# thresh <- 10
# ppd <- ppd %>%
#   ungroup() %>%
#   mutate(removed = ifelse(n < thresh, T, F))
# 
# # what proportion of vulture*days are removed overall?
# (propRemoved <- ppd %>%
#     group_by(seasonUnique, season) %>%
#     summarize(nTot = n(),
#               nRemoved = sum(removed),
#               propRemoved = nRemoved/nTot))
# propRemoved %>%
#   ggplot(aes(x = seasonUnique, y = propRemoved, col = season))+
#   geom_point(size = 10)+
#   theme_classic()+
#   scale_color_manual(values = c("blue", "red", "darkorange"))
# 
# # I'm a bit concerned about this because it removes more data for the breeding season than for the non-breeding seasons (likely because less sunlight during those seasons)
# 
# # what about by individual over each season?
# ppd %>%
#   group_by(seasonUnique, season, Nili_id) %>%
#   summarize(propRemoved = sum(removed)/n()) %>%
#   ggplot(aes(x = seasonUnique, y = propRemoved))+
#   geom_boxplot(aes(group = seasonUnique, fill = season))+
#   geom_jitter(width = 0.2, alpha = 0.3, size = 2)+
#   scale_fill_manual(name = "Season", values = c("blue", "red", "darkorange"))+
#   theme_classic()+ # tells the same story
#   ylab("Proportion of days removed")+
#   xlab("")

## Remove days on which individuals don't have at least 10 points.
before <- seasons
seasons <- map(seasons, ~.x %>%
                 group_by(Nili_id, dateOnly) %>%
                 filter(n() >= 10) %>% # hard-coding the 10 in here so it doesn't accidentally get changed.
                 ungroup())
after <- seasons

beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows) %>%
  mutate(propRowsRemoved = round((nBefore-nAfter)/nBefore, 2))
```

Here's how many points get removed when we do this filtering:

```{r include = T, echo = F}
df
```

### Step 11: Remove individuals based on number of days tracked

This filtering comes after the removal of some vulture\*days for having
too few points. So, considering the remaining days, we only want to
retain individuals that are tracked for at least 1/3 of a given season
(in terms of number of days).

Here are the number of days in each season. As you can see, breeding
season is the longest, followed by summer, then fall. (Ignore Summer
2020--it will not be included in the analysis.)

```{r include = TRUE, echo = F}
seasonNames
durs <- map_dbl(seasons, ~length(unique(.x$dateOnly)))
durs
```

```{r remove-daysTracked}
# Include only individuals with enough days tracked -----------------------
## Must be tracked for at least 1/3 of the number of days in the season.

# walk2(seasons, durs, ~.x %>%
#   st_drop_geometry() %>%
#   group_by(Nili_id) %>%
#   summarize(nDaysTracked = length(unique(dateOnly)),
#             propDaysTracked = nDaysTracked/.y) %>%
#   pull(propDaysTracked) %>%
#   hist())

indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>%
                       st_drop_geometry() %>%
                       group_by(Nili_id) %>%
                       summarize(nDaysTracked = length(unique(dateOnly)),
                                 propDaysTracked = nDaysTracked/.y) %>%
                       filter(propDaysTracked >= 0.33) %>% # XXX ask Orr if this should be percentage vs. number of days, given the different season lengths. (email)
                       pull(Nili_id))

## remove those individuals not tracked for enough days
before <- seasons
after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y))

beforeRows <- map_dbl(before, nrow)
afterRows <- map_dbl(after, nrow)
df <- data.frame(season = seasonNames, nBefore = beforeRows, nAfter = afterRows) %>%
  mutate(propRowsRemoved = round((nBefore-nAfter)/nBefore, 2))

```

Here's what this does to the row counts:

```{r include = T, echo = F}
df
```

### Step 12: Calculate flight altitudes

Using the `elevatr` package, I get ground elevation values for each gps
point in each season, at zoom level 10 (which should come out to around
150 meter cell resolution in the resulting raster, according to )

```{r get-elevation}
# Get elevation rasters ---------------------------------------------------
elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x %>% st_as_sf(coords = c("location_long", "location_lat"), remove = FALSE, crs = "WGS84"), z = 10))
# 
save(elevs_z10_seasons, file = "data/elevs_z10_seasons.Rda")
load("data/elevs_z10_seasons.Rda")

groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y %>% sf::st_as_sf(coords = c("location_long", "location_lat"), crs = "WGS84", remove = FALSE))) # have to convert the crs because the raster from elevatr is in WGS84.

## use elevation rasters to calculate height above ground level
seasons <- map2(.x = seasons, .y = groundElev_z10, 
                ~.x %>% mutate(height_above_ground = .y,
                               height_above_ground = case_when(height_above_ground <0 ~ 0,
                                                               TRUE ~ height_above_ground)))

# Export the data
save(seasons, file = "data/seasons.Rda")
```

As expected, adding the elevation information doesn't change the number
of rows in the data.

### Step 13: Downsample the data

Because occasionally, the vultures' tags were set to read more
frequently than every 10 minutes, I want to downsample the data to a
10-minute interval.

```{r downsample}
# Downsample the data and save the downsampled data
subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){
  sub <- df %>%
    arrange(idCol, timestampCol) %>%
    group_by(.data[[idCol]]) %>%
    mutate(tc = cut(.data[[timestampCol]], 
                    breaks = paste(as.character(mins), "min", sep = " "))) %>%
    group_by(.data[[idCol]], 
             "d" = lubridate::date(.data[[timestampCol]]), tc) %>%
    slice(1) %>%
    ungroup() %>%
    dplyr::select(-c("d", "tc"))
  return(sub)
}

# create and save downsampled datasets for later use
seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T)
seasons_20min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 20), .progress = T)
seasons_30min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 30), .progress = T)
seasons_60min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 60), .progress = T)
seasons_120min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 120), .progress = T)

map_dbl(seasons, nrow)
map_dbl(seasons_10min, nrow)

# save(seasons_10min, file = "data/seasons_10min.Rda")
# save(seasons_20min, file = "data/seasons_20min.Rda")
# save(seasons_30min, file = "data/seasons_30min.Rda")
# save(seasons_60min, file = "data/seasons_60min.Rda")
# save(seasons_120min, file = "data/seasons_120min.Rda")

# no need to calculate roosts for the downsampled data because in theory there's only one point per night.

df <- data.frame(season = seasonNames, 
                 nBefore = map_dbl(seasons, nrow),
                 nAfter = map_dbl(seasons_10min, nrow)) %>%
  mutate(propChange = round((nBefore-nAfter)/nBefore, 2))
```

Here are the row counts after downsampling:

```{r include = TRUE, echo = F}
df
```

And here are the final counts of number of vultures:

```{r final-counts}
df <- seasons_10min %>% purrr::list_rbind() %>% 
  group_by(seasonUnique) %>% 
  summarize(vultures = length(unique(Nili_id)))
```

```{r include = TRUE, echo = F}
df
```