### 2023-08-17 00_dataPrepReport.Rmd
progress on the data prep report as rmarkdown doc

---
title: "Data preparation report"
author: "Kaija Gahm"
date: "2023-05-22"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE, echo = FALSE, message = FALSE, cache = FALSE)
```

This is a condensed and simplified data preparation report. It excludes some exploratory plotting and analyses. There are records of those in 00_dataPrep.R.

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
library(elevatr)
```

## 1. Download data from Movebank
```{r}
# ## Download data from movebank
# base::load("movebankCredentials/pw.Rda")
# MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
# rm(pw)
# 
# # Hatzofe INPA data
# # inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
# # inpadf <- methods::as(inpa, "data.frame")
# # inpadf_touse <- inpadf %>%
# #   mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
# #          year = as.numeric(lubridate::year(timestamp))) %>%
# #   filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-05-15")) # cut this off at the same point as the ornitela data
# # write_feather(inpadf_touse, "data/inpadf_touse.feather")
inpadf_touse <- read_feather("data/inpadf_touse.feather")
# # Ornitela: download data from movebank (just a subset of the times for now)
# minDate <- "2020-09-01 00:00"
# maxDate <- "2023-05-15 11:59"
# # dat <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
# # write_feather(dat, "data/dat.feather")
dat <- read_feather("data/dat.feather")
```

When we download the data, there are `r (a1 <- nrow(dat))` GPS fixes from the Ornitela dataset and `r (a2 <- nrow(inpadf_touse))` fixes from the INPA dataset. Their date ranges are from `r (a3 <- min(lubridate::date(dat$timestamp), na.rm = T))` to `r (a4 <- max(lubridate::date(dat$timestamp), na.rm = T))` (Ornitela) and `r (a5 <- min(lubridate::date(inpadf_touse$timestamp), na.rm = T))` to `r (a6 <- max(lubridate::date(inpadf_touse$timestamp), na.rm = T))` (INPA). The Ornitela data contains `r (a7 <- length(unique(dat$trackId)))` individuals, and the INPA dataset contains `r (a8 <- length(unique(inpadf_touse$trackId)))`, for a total of `r (a9 <- length(unique(c(dat$trackId, inpadf_touse$trackId))))` unique birds.

Products: **dat** (ornitela data); **inpadf_touse** (inpa data, prepared to join)
Metrics: how many individuals in each, how many rows in each

## 2. Join INPA to Ornitela data. Clean ID's.
```{r}
# Will there be a problem joining these?
names(dat)[!(names(dat) %in% names(inpadf_touse))] # all names in "dat" are found in inpadf_touse
names(inpadf_touse)[!(names(inpadf_touse) %in% names(dat))] # many are not found in "dat" that are found in inpadf_touse.

# Add the dataset names so we can keep track of where the data comes from
inpadf_tojoin <- inpadf_touse[,names(dat)] %>%
  mutate(dataset = "inpa")
dat <- dat %>%
  mutate(dataset = "ornitela")

# Join the two datasets
joined <- bind_rows(inpadf_tojoin, dat)
dim(joined)

## fix trackId
dat2 <- joined %>%
  mutate(trackId = as.character(trackId),
         trackId = case_when(trackId == "E03" ~ "E03w",
                             TRUE ~ trackId))
# number of unique individuals
length(unique(dat2$trackId)) # 206

# Add the Nili_id
ww <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, Movebank_id) %>%
  distinct()

# Are there any other individuals who have two different associated trackId's?
ww %>% filter(Movebank_id %in% dat2$trackId) %>% dplyr::select(Nili_id, Movebank_id) %>% distinct() %>% group_by(Nili_id) %>% summarize(n = length(unique(Movebank_id))) %>% arrange(desc(n)) # Note that there is one Nili_id in our dataset that has two associated trackId's--it's yomtov.
ww %>% filter(Nili_id =="yomtov") # Y26 and Y26b. So we should expect to once again "lose" an individual once we change to Nili ID's

all(dat2$trackId %in% ww$Movebank_id) # false.

# Okay, let's investigate the ones that aren't included
dat2 %>%
  filter(!(trackId %in% ww$Movebank_id)) %>%
  pull(trackId) %>% unique() # looks like this is going to be an issue of character matching. Let's transform everything to lowercase and remove all characters and spaces

dat2$trackId <- str_replace_all(tolower(dat2$trackId), "\\.", "")
ww$mb_id <- str_replace_all(tolower(ww$Movebank_id), "\\s|\\(|\\)|\\>", "") %>% str_replace_all(., "whiite", "white")

sort(unique(ww$mb_id))
unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
dat2$trackId[dat2$trackId == "y01t60w"] <- "t60white"
dat2$trackId[dat2$trackId == "e86white"] <- "e86"
unique(dat2$trackId)[!(unique(dat2$trackId) %in% ww$mb_id)]
# Investigate the remaining ones:
dat2 %>% filter(!(trackId %in% ww$mb_id)) %>% dplyr::select(dateOnly, trackId, local_identifier, ring_id, dataset) %>% group_by(trackId, local_identifier, ring_id, dataset) %>% summarize(earlyDate = min(dateOnly)) # okay, looks like these three come from the INPA dataset and haven't been entered into the who's who yet.
# I guess for now let's just make their trackId's into Nili_id's

dat3 <- left_join(dat2, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
                  by = c("trackId" = "mb_id"))
nrow(dat3) == nrow(dat2) # good, same number of rows. (check that the join worked)
length(unique(dat3$Nili_id))
# For any that didn't get Nili_ids assigned, make it the trackId
dat3 %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expecfted.
dat3 <- dat3 %>%
  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
                             TRUE ~ Nili_id))
unique(dat3$Nili_id)
```

After joining the two datasets together and cleaning up the names, we end up with a single dataset with `r (b1 <- length(unique(dat3$Nili_id)))` individuals and `r (b2 <- nrow(dat3))` fixes, and with a date range from `r (b3 <- min(lubridate::date(dat3$timestamp), na.rm = T))` to `r (b4 <- max(lubridate::date(dat3$timestamp), na.rm = T))`.
Product: **dat3**
Metrics: number of rows; number of individuals; date range of individuals

## 3. Remove periods in hospital/cages
```{r}
# ## annotate the data with periods to remove
toRemove <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "periods_to_remove")

toRemove <- toRemove %>%
  dplyr::select(Nili_id,
                remove_start,
                remove_end,
                reason) %>%
  mutate(across(c(remove_start, remove_end), .fns = function(x){
    lubridate::ymd(x)
  })) %>%
  dplyr::filter(!is.na(remove_end))

toRemove_long <- toRemove %>%
  group_by(Nili_id, reason) %>%
  # sequence of daily dates for each corresponding start, end elements
  dplyr::mutate(dateOnly = map2(remove_start, remove_end, seq, by = "1 day")) %>%
  # unnest the list column
  unnest(cols = c(dateOnly)) %>%
  # remove any duplicate rows
  distinct() %>%
  dplyr::select(-c(remove_start, remove_end)) %>%
  rename("status" = reason)

# Join to the original data
datAnnot <- dat3 %>%
  left_join(toRemove_long, by = c("Nili_id", "dateOnly")) %>%
  mutate(status = replace_na(status, "valid"))
nrow(datAnnot) == nrow(dat3) #T: good, same number of rows.

# Actually REMOVE the periods to remove...
datAnnot <- datAnnot %>%
  filter(status == "valid")
nrow(datAnnot)
length(unique(datAnnot$Nili_id))
nrow(dat3)
nrow(datAnnot)

#
# Identify and remove capture dates using Marta's code --------------------
## To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function). Then, if the bird roosted within 500 m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset. This has been validated. Note: this protocol doesn't work for the Carmel because of the position of the roost sites relative to cages. But we're just dealing with southern individuals, so that's okay.
# roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
# save(roosts, file = "data/roosts.Rda")
load("data/roosts.Rda")

# Identify the period of time during which the capture sites are open (when we need to do this exclusion)
start.day <- 01
start.month <- 08
end.day <-  30
end.month <- 11
distance <- 500 # distance, in meters, to calculate from the cage

# Load the information about the capture sites
sites <- read.csv("data/capture_sites.csv")

# Subset the roost dataset with the start and end dates for the capture period
sub.roosts <- roosts %>%
  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
                              format="%d-%m-%Y"),
         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
                            format="%d-%m-%Y")) %>%
  filter(date >= start_date & date <= end_date)

unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.

# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat),
               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates

DistanceMat <- matrix(ncol = nrow(sites), nrow = nrow(crds))
colnames(DistanceMat) <- unique(sites$name)

for(i in 1:nrow(crds)){ # for each roost point...
  DistanceMat[i,] <- round(geosphere::distm(crds[i,], sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
  #print(i)
}

ClosestCaptureSite <- colnames(DistanceMat)[apply(DistanceMat,1,which.min)] # ID of closest capture site
ClosestCaptureDist <- apply(DistanceMat,1,min) # Distance from closest capture site

sub.roosts <- cbind(sub.roosts, ClosestCaptureSite, ClosestCaptureDist)
sub.roosts$Captured <- ifelse(sub.roosts$ClosestCaptureDist <= distance, "Yes", "No")

sub.captured <- subset(sub.roosts, Captured == "Yes") # get list of dates which the bird was inside the cage

sub.captured.dates <- subset(sub.captured,
                             select = c("Nili_id",
                                        "date",
                                        "ClosestCaptureSite",
                                        "ClosestCaptureDist",
                                        "Captured"))

## This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it. So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data

sub.captured.no.carmel <- subset(sub.captured.dates, ClosestCaptureSite != "Carmel")
sub.captured.carmel <- subset(sub.captured.dates, ClosestCaptureSite == "Carmel")

AllCarmelDates <- read.csv("data/all_captures_carmel_2010-2021.csv")
AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")

AllCarmelDates.1 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-1)))
AllCarmelDates.2 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-2)))
AllCarmelDates.3 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-3)))

AllCarmelDates.all <- rbind(AllCarmelDates, AllCarmelDates.1, AllCarmelDates.2, AllCarmelDates.3)

sub.captured.carmel <- sub.captured.carmel %>%
  mutate(known_capture = ifelse(date %in% AllCarmelDates.all$Date, 1, 0),
         Captured = ifelse(known_capture == 1 & ClosestCaptureDist <= 50, "yes", "no")) %>%
  filter(Captured == "yes") %>%
  dplyr::select(-c(known_capture))

names(sub.captured.no.carmel)
names(sub.captured.carmel)

sub.captured.dates <- rbind(sub.captured.no.carmel, sub.captured.carmel)

# We also need to exclude the day after the bird was captured
sub.captured.dates.1 <- sub.captured.dates
sub.captured.dates.1$date <- sub.captured.dates.1$date+1

sub.captured.dates <- rbind(sub.captured.dates, sub.captured.dates.1)
sub.captured.dates <- sub.captured.dates %>%
  dplyr::distinct(Nili_id, date, .keep_all = T)

# It all looks ok, so we can subset the dataset to exclude the capture dates
datAnnot_noCaptures <- datAnnot %>%
  left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
nrow(datAnnot) == nrow(datAnnot_noCaptures) # should be TRUE. NOW we can filter.
datAnnot_noCaptures <- datAnnot_noCaptures %>%
  dplyr::filter(Captured != "Yes"|is.na(Captured)) # remove the individual*days when they were captured

# How many rows and individuals did we remove?
before <- nrow(datAnnot)
after <- nrow(datAnnot_noCaptures)
(propChange <- (before-after)/before) # approx. 1% of data removed.
length(unique(datAnnot$Nili_id)) == length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.

datAnnot_noCaptures <- datAnnot_noCaptures %>%
  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))
```

After removing periods when individuals were in the hospital, the number of fixes has been reduced to `r (c1 <- nrow(datAnnot))`, which is a `r `(round(100*(b2-c1)/b2, 3))`% decrease. The number of individuals has not changed: it still has `r (c2 <- length(unique(datAnnot$Nili_id)))` unique vultures.

Then we remove periods of time when individuals were in the capture cages. We are left with `r (c3 <- nrow(datAnnot_noCaptures))` fixes (a further `r `(round(100*(c1-c3)/c1, 3))`% decrease) and `r (c4 <- length(unique(datAnnot_noCaptures$Nili_id)))` individuals.

Product: **datAnnot** (after removing periodsToRemove/hospital stays), datAnnot_noCaptures (after removing capture cage periods)
Metrics: number of rows

## 4. Attach age and sex
```{r}
# Age and sex info --------------------------------------------------------
# Attach age and sex information
as <- read_excel("data/whoswho_vultures_20230315_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, birth_year, sex) %>%
  distinct()

datAnnot2 <- datAnnot_noCaptures %>%
  dplyr::select(-c("sex")) %>%
  left_join(as, by = "Nili_id")
nrow(datAnnot2) == nrow(datAnnot)
length(unique(datAnnot2$Nili_id)) # 196

datAnnot <- datAnnot2 # a hack so the rest of the code will work
```

After attaching the age and sex information, the number of rows and individuals has not changed (`r (d1 <- nrow(datAnnot2))` rows, `r (d2 <- length(unique(datAnnot2$Nili_id)))` individuals).

Product: **datAnnot2** (then it gets renamed to datAnnot so the rest of the code will run, but datAnnot2 is a direct parallel of the state of datAnnot at the end of this).
Metrics: number of columns before and after; number of rows (to show that none were added)

## 5. Clean data (remove invalid points; keep individuals with at least 30 days total in the entire dataset)
```{r}
# Clean the data
## Region masking, downsampling, removal of speed outliers, setting altitude outliers to NA, etc.
mask <- sf::st_read("data/CutOffRegion.kml")
#datAnnotCleaned <- vultureUtils::cleanData(dataset = datAnnot, mask = mask, inMaskThreshold = 30, removeVars = F, idCol = "Nili_id", downsample = F, reMask = T) # XXX using 30 days overall. If we want 30 days per season, will have to do that differently (e.g. split the seasons earlier)--but I think that that'll be covered later in the data cleaning anyway.
#save(datAnnotCleaned, file = "data/datAnnotCleaned.Rda")
load("data/datAnnotCleaned.Rda")
length(unique(datAnnotCleaned$Nili_id)) # 172 individuals
# Fix time zone so dates make sense ---------------------------------------
## Overwrite the dateOnly column from the new times
datAnnotCleaned <- datAnnotCleaned %>%
  mutate(timestampIsrael = lubridate::with_tz(timestamp, tzone = "Israel"),
         dateOnly = lubridate::date(timestampIsrael),
         month = lubridate::month(dateOnly),
         year = lubridate::year(dateOnly))
```

Next, we clean the data with the `vultureUtils::cleanData()` function, removing invalid points. We retain only vultures that are tracked for at least 30 days in the entire dataset. We are left with `r (e1 <- nrow(datAnnotCleaned))` fixes from `r (e2 <- length(unique(datAnnotCleaned$Nili_id)))` unique vultures. This is a `r round(100*((d1-e1)/d1), 3)`% decrease in fixes and a `r (round(100*((d2-e2)/d2), 3))`% decrease in individuals.

Product: **datAnnotCleaned**
Metrics: number of individuals, number of rows

## 6. Split into seasons
```{r}
# # Split into Marta's 3 seasons --------------------------------------------
# s1 <- datAnnotCleaned %>%
#   mutate(start_breeding = ifelse(month %in% c(1:5),
#                                  paste(as.character(year-1), "-", "12", "-", "15", sep = ""),
#                                  paste(as.character(year), "-", "12", "-", "15", sep = "")),
#          start_summer = ifelse(month == 12,
#                                paste(as.character(year + 1), "-", "05", "-", "15", sep = ""),
#                                paste(as.character(year), "-", "05", "-", "15", sep = "")),
#          start_fall = as.Date(paste(as.character(year), "-", "09", "-", "15", sep = "")),
#          start_breeding = as.Date(start_breeding),
#          start_summer = as.Date(start_summer),
#          season = case_when(
#            dateOnly >= start_breeding & dateOnly < start_summer ~ "breeding",
#            dateOnly >= start_summer & dateOnly <= start_fall ~ "summer",
#            dateOnly >= start_fall & dateOnly < start_breeding ~ "fall")) %>%
#   filter(dateOnly != "2023-05-15") # remove May 15 2023, because it turns out the season boundaries are supposed to be non-inclusive and this is the latest one.
# 
# s2 <- s1 %>%
#   mutate(age = year - birth_year,
#          sex = case_when(sex == "f/m" ~ "u",
#                          TRUE ~ sex)) %>%
#   mutate(age = ifelse(season == "breeding" & month %in% c(1, 12), age + 1, age)) %>%
#   group_by(Nili_id, year, month) %>%
#   mutate(age = ifelse(month == 2, min(age) + 1, age)) %>%
#   mutate(age_group = case_when(age == 0 ~ "juv",
#                                age >= 1 & age <= 4 ~ "sub",
#                                age >= 5 ~ "adult",
#                                TRUE ~ NA)) %>%
#   mutate(seasonUnique = case_when(season %in% c("fall", "summer") ~ paste(year, season, sep = "_"),
#                                   season == "breeding" & month == 12 ~ paste(year + 1, season, sep = "_"),
#                                   season == "breeding" & month != 12 ~ paste(year, season, sep = "_"))) %>%
#   # explicitly set the factor levels in temporal order
#   mutate(seasonUnique = factor(seasonUnique, levels = c("2020_summer", "2020_fall", "2021_breeding", "2021_summer", "2021_fall", "2022_breeding", "2022_summer", "2022_fall", "2023_breeding")),
#          season = factor(season, levels = c("breeding", "summer", "fall")))
# table(s2$seasonUnique, exclude = NULL) # XXX can use this in the report--easy to read
# 
# # remove columns that were used in the season calculation
# s3 <- s2 %>%
#   dplyr::select(-c(month, start_breeding, start_summer, start_fall))
# 
# # Separate the seasons -----------------------------
# seasons_orig <- s3 %>%
#   group_by(seasonUnique) %>%
#   group_split(.keep = T)
# # extract the season names in case we need a separate vector at some point.
# seasonNames_orig <- map_chr(seasons_orig, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
# save(seasons_orig, file = "data/seasons_orig.Rda")
# save(seasonNames_orig, file = "data/seasonNames_orig.Rda")
load("data/seasons_orig.Rda")
load("data/seasonNames_orig.Rda")
```

We've split the data into seasons now. The season names are: `r (f1 <- seasonNames_orig)`. We have a total of `r (f2 <- length(f1))` seasons. Here are the date ranges, numbers of fixes, and numbers of individuals for each season:
```{r echo = F, include = T}
mins <- map_chr(seasons_orig, ~as.character(min(lubridate::date(.x$timestamp), na.rm = T)))
maxes <- map_chr(seasons_orig, ~as.character(max(lubridate::date(.x$timestamp), na.rm = T)))
fixes <- map_dbl(seasons_orig, nrow)
inds <- map_dbl(seasons_orig, ~length(unique(.x$Nili_id)))
f3 <- data.frame(season = seasonNames_orig, startDate = mins, endDate = maxes, nFixes = fixes, nVultures = inds)
f3
```

Products: seasons_orig, seasonNames_orig
Metrics: number of seasons; number of individuals and rows per season; which seasons are included; date ranges for each season.

## 7. Remove individuals with too low a fix rate: Exclude individuals that never have a daily mode of 10 minutes
```{r}
# Optionally remove individuals with a low fix rate -----------------------
Mode <- function(x) { # function to calculate mode fix rate
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Calculate daily modes and add them to the dataset
seasons <- map(seasons_orig, ~.x %>%
                 group_by(dateOnly, Nili_id) %>%
                 mutate(diff = as.numeric(difftime(lead(timestamp), timestamp, units = "mins"))) %>%
                 group_by(dateOnly, Nili_id) %>%
                 mutate(mode = Mode(round(diff))) %>%
                 ungroup())

# Exclude individuals that never have a daily mode of 10 minutes
toKeep_fixrate <- map(seasons, ~.x %>%
                sf::st_drop_geometry() %>%
                group_by(Nili_id) %>%
                summarize(minmode = min(mode, na.rm = T)) %>%
                filter(minmode <= 10) %>%
                pull(Nili_id) %>%
                unique())
save(toKeep_fixrate, file = "data/toKeep_fixrate.Rda")
load("data/toKeep_fixrate.Rda")
seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y))
```
Product: seasons, seasons_mode10
Metrics: number of individuals per season, number of fixes per season

After removing individuals that never have a 10-minute fix rate, here are the numbers of fixes and numbers of individuals for each season:
```{r echo = F, include = T}
fixes <- map_dbl(seasons_mode10, nrow)
inds <- map_dbl(seasons_mode10, ~length(unique(.x$Nili_id)))
g1 <- data.frame(season = seasonNames_orig, nFixes = fixes, nVultures = inds)
g1
```

At this point, we save the data for social analysis, saving both seasons_forSoc (which contains all individuals, regardless of fix rate) and seasons_forSoc_mode10 (which contains only individuals with a 10-minute fix rate).

## 8. Save SOCIAL data
```{r}
# Save copies for social analysis
# The next step will be to remove individuals with too few points per day or too few days tracked. But we don't want those indivs removed for the *social* analysis, since those things will be accounted for with SRI and all individuals make up important parts of the social fabric. So, before I filter for ppd and for days tracked, going to save a copy to use for social analysis.
seasons_forSoc <- seasons
seasons_forSoc_mode10 <- seasons_mode10
save(seasons_forSoc, file = "data/seasons_forSoc.Rda")
save(seasons_forSoc_mode10, file = "data/seasons_forSoc_mode10.Rda")
load("data/seasons_forSoc.Rda")
load("data/seasons_forSoc_mode10.Rda")
# rename, preparing to carry on
seasons_mode10 <- seasons_forSoc_mode10
seasons <- seasons_forSoc

# Investigate: How many individuals are retained in each of these cases? What proportion were removed?
inds_all <- map_dbl(seasons_forSoc, ~length(unique(.x$Nili_id)))
inds_mode10 <- map_dbl(seasons_forSoc_mode10, ~length(unique(.x$Nili_id)))
propRemoved <- round((inds_all-inds_mode10)/inds_all, 3) # For all seasons, getting rid of individuals that don't have
propRemoved # yikes that's a lot of birds. Alas.
```

Here are the numbers of individuals for each of these products:
```{r include = T, echo = F}
inds_reg <- map_dbl(seasons_forSoc, ~length(unique(.x$Nili_id)))
inds_mode10 <- map_dbl(seasons_forSoc_mode10, ~length(unique(.x$Nili_id)))
h1 <- data.frame(season = seasonNames_orig, allIndivs = inds_reg, highFixRate = inds_mode10) %>% mutate(propHighFixRate = round(highFixRate/allIndivs, 2))
h1
```

Products: seasons_forSoc, seasons_forSoc_mode10. These shouldn't be any different from seasons and seasons_mode10, there's just some renaming that happens.

## 9. Get ROOSTS
```{r}
# Get roosts for each season ----------------------------------------------
# roosts_seasons <- purrr::map(seasons, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
# roosts_seasons <- roosts_seasons %>%
#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
# save(roosts_seasons, file = "data/roosts_seasons.Rda")
load("data/roosts_seasons.Rda")

# roosts_seasons_mode10 <- purrr::map(seasons_mode10, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
# roosts_seasons_mode10 <- roosts_seasons_mode10 %>%
#   map(., ~st_as_sf(.x, crs = "WGS84", coords = c("location_long", "location_lat"), remove = F))
# save(roosts_seasons_mode10, file = "data/roosts_seasons_mode10.Rda")
load("data/roosts_seasons_mode10.Rda")
```
Products: roosts_seasons and roosts_seasons_mode10
Metrics: number of individuals per season, number of rows per season

Now, we use the full dataset (not restricted to high fix rate individuals) to obtain a locations for each individual on each night. This gives us:
```{r echo = F, include = T}
cat("All individuals:\n")
(i1 <- data.frame(season = seasonNames_orig, 
                 nIndivs = map_dbl(roosts_seasons, ~length(unique(.x$Nili_id))),
                 vultureNights = map_dbl(roosts_seasons, nrow)))
```

And for comparison, the vultures and vultureNights for roosts including only the high fix rate individuals:

```{r echo = F, include = T}
cat("\nHigh fix rate individuals:\n")
(i2 <- data.frame(season = seasonNames_orig, 
                 nIndivs = map_dbl(roosts_seasons_mode10, ~length(unique(.x$Nili_id))),
                 vultureNights = map_dbl(roosts_seasons_mode10, nrow)))
```

## 10. Remove night points
```{r}
# Remove nighttime points -------------------------------------------------
before <- map_dbl(seasons, nrow)
beforeindivs <- map_dbl(seasons, ~length(unique(.x$Nili_id)))
seasons <- map(seasons, ~{
  times <- suncalc::getSunlightTimes(date = unique(lubridate::date(.x$timestamp)),
                                     lat = 31.434306, lon = 34.991889, # XXX revisit this!
                                     keep = c("sunrise", "sunset")) %>%
    dplyr::select("dateOnly" = date, sunrise, sunset)

  .x <- .x %>%
    dplyr::mutate(dateOnly = lubridate::ymd(dateOnly)) %>%
    dplyr::left_join(times, by = "dateOnly") %>%
    dplyr::mutate(daylight = ifelse(timestamp >= sunrise & timestamp <= sunset, "day", "night")) %>%
    dplyr::select(-c(sunrise, sunset)) %>%
    dplyr::filter(daylight == "day")
})
# after <- map_dbl(seasons, nrow)
# change <- after-before
# changeDF <- data.frame(seasonName = seasonNames_orig, nBefore = before, nAfter = after) %>%
#   mutate(propRowsRemoved = round((before-after)/before, 2))
# changeDF
```
Product: seasons/"after" (maybe need a unique name here?)
Metrics: number of individuals per season (should be the same), number of rows per season

After removing the night points, the number of individuals per season should not have changed, but some rows have been removed.
```{r}
j1 <- data.frame(season = seasonNames_orig, 
                 nIndivsBefore = beforeindivs,
                 nIndivsNow = map_dbl(seasons, ~length(unique(.x$Nili_id))),
                 nIndivs = map_dbl(roosts_seasons, ~length(unique(.x$Nili_id))),
                 vultureNights = map_dbl(roosts_seasons, nrow))
```

<!-- ## 11. Restrict to southern individuals -->
<!-- ```{r} -->
<!-- # Restrict to southern individuals ---------------------------------------- -->
<!-- # Based on previous investigations for the 2022 breeding and non-breeding seasons, have found that a good cutoff for southern vs. non-southern is 3550000 (UTM 36N, https://epsg.io/32636) -->
<!-- ## Transform to SF object, so we can get centroids -->
<!-- # seasonsSF <- map(seasons_orig, ~.x %>% -->
<!-- #                    sf::st_as_sf(coords = c("location_long", "location_lat"), remove = F) %>% -->
<!-- #                    sf::st_set_crs("WGS84") %>% -->
<!-- #                    sf::st_transform(32636)) -->
<!-- #  -->
<!-- # save(seasonsSF, file = "data/seasonsSF.Rda") -->
<!-- load("data/seasonsSF.Rda") -->

<!-- ## Get centroids, so we can see who's "southern" for that season. -->
<!-- centroids <- map(seasonsSF, ~.x %>% -->
<!--                    group_by(Nili_id) %>% -->
<!--                    summarize(geometry = st_union(geometry)) %>% -->
<!--                    st_centroid()) -->

<!-- ## Examine a histogram of centroid latitudes  -->
<!-- # walk(centroids, ~hist(st_coordinates(.x)[,2])) # looks like 3550000 is generally a good cutoff point here. -->

<!-- ## Get southern individuals for each season, so we can filter the data -->
<!-- southernIndivs <- map(centroids, ~.x %>% -->
<!--                         filter(st_coordinates(.)[,2] < 3550000) %>% -->
<!--                         pull(Nili_id)) -->

<!-- ## Remove individuals not in the south -->
<!-- before <- seasons -->
<!-- seasons <- map2(.x = seasons, .y = southernIndivs, ~.x %>% filter(Nili_id %in% .y)) -->
<!-- after <- seasons -->

<!-- beforeRows <- map_dbl(before, nrow) -->
<!-- afterRows <- map_dbl(after, nrow) -->
<!-- beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id))) -->
<!-- afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id))) -->
<!-- df <- data.frame(season = seasonNames_orig, nBefore = beforeRows, nAfter = afterRows, indivsBefore = beforeIndivs, indivsAfter = afterIndivs) -->
<!-- df$propRowsRemoved <- round((df$nBefore - df$nAfter)/df$nBefore, 2) -->
<!-- df -->
<!-- rm(centroids) -->
<!-- rm(seasonsSF) -->
<!-- ``` -->
<!-- Product: seasons/after (unique name?) -->
<!-- Metrics: number of individuals per season, number of rows per season, boxplot of days per individual per season (so we can compare?) -->

<!-- ## 12. Include only days with > 10ppd unless the battery never dropped below 50% on that day -->
<!-- ```{r} -->
<!-- ## Keep days with more than 10 points, or with fewer than 10 if the minimum battery charge that day is >50% -->
<!-- seasons <- map(seasons, ~.x %>% -->
<!--                  group_by(Nili_id, dateOnly) %>% -->
<!--                  filter(n() >= 10 | (n() < 10 & min(battery_charge_percent, na.rm = T) > 50)) %>% -->
<!--                  ungroup()) -->
<!-- ``` -->
<!-- Product: seasons (unique name?) -->
<!-- Metrics: number of individuals per season (should not have changed since previous), number of rows per season, boxplot of days per individual per season (compare to previous) -->

<!-- ## 13. For each season, exclude individuals with < 30 valid days (after removing days with too few points) -->
<!-- ```{r} -->
<!-- # Include only individuals with enough days tracked ----------------------- -->
<!-- ## Must be tracked for at least 30 days per season -->
<!-- durs <- map_dbl(seasons, ~length(unique(.x$dateOnly))) -->

<!-- # walk2(seasons, durs, ~.x %>% -->
<!-- #         st_drop_geometry() %>% -->
<!-- #         group_by(Nili_id) %>% -->
<!-- #         summarize(nDaysTracked = length(unique(dateOnly)), -->
<!-- #                   propDaysTracked = nDaysTracked/.y) %>% -->
<!-- #         pull(propDaysTracked) %>% -->
<!-- #         hist()) -->

<!-- indivsToKeep <- map2(.x = seasons, .y = durs, ~.x %>% -->
<!--                        st_drop_geometry() %>% -->
<!--                        group_by(Nili_id) %>% -->
<!--                        summarize(nDaysTracked = length(unique(dateOnly)), -->
<!--                                  propDaysTracked = nDaysTracked/.y) %>% -->
<!--                        filter(nDaysTracked >= 30) %>% -->
<!--                        pull(Nili_id)) -->

<!-- ## remove those individuals not tracked for enough days -->
<!-- before <- seasons -->
<!-- after <- map2(.x = seasons, .y = indivsToKeep, ~.x %>% filter(Nili_id %in% .y)) -->

<!-- beforeIndivs <- map_dbl(before, ~length(unique(.x$Nili_id))) -->
<!-- afterIndivs <- map_dbl(after, ~length(unique(.x$Nili_id))) -->
<!-- seasons <- after -->
<!-- ``` -->
<!-- Product: seasons/after (unique name) -->
<!-- Metric: number of individuals per season (should have decreased), number of points per season, boxplot of days per individual per season -->

<!-- ## 14. Get elevation information (and remove summer 2020 because it has too little data) -->
<!-- ```{r} -->
<!-- # Get elevation rasters --------------------------------------------------- -->
<!-- # Remove the first element of `seasons` because it doesn't have any data -->
<!-- # seasons <- seasons[-1] # removing summer 2020 -->
<!-- # # XXXXXXX start here -->
<!-- # seasons <- map(seasons, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84")) -->
<!-- # elevs_z10_seasons <- map(seasons, ~elevatr::get_elev_raster(.x , z = 10)) -->
<!-- # save(elevs_z10_seasons, file = "data/elevs_z10_seasons.Rda") -->
<!-- load("data/elevs_z10_seasons.Rda") -->

<!-- # groundElev_z10 <- map2(elevs_z10_seasons, seasons, ~raster::extract(x = .x, y = .y)) # have to convert the crs because the raster from elevatr is in WGS84. -->
<!-- # save(groundElev_z10, file = "data/groundElev_z10.Rda") -->
<!-- load("data/groundElev_z10.Rda") -->

<!-- ## use elevation rasters to calculate height above ground level -->
<!-- before <- seasons -->
<!-- seasons <- map2(.x = seasons, .y = groundElev_z10,  -->
<!--                 ~.x %>% mutate(height_above_ground = .y, -->
<!--                                height_above_ground = case_when(height_above_ground < 0 ~ 0, -->
<!--                                                                TRUE ~ height_above_ground))) -->
<!-- after <- seasons -->

<!-- # There should be no change in the number of rows -->
<!-- all(map2_lgl(before, after, ~nrow(.x) == nrow(.y))) -->
<!-- ``` -->
<!-- Product: seasons/after -->
<!-- Metrics: number of seasons (should be one fewer than before), number of rows per season (expect no change), number of columns, histograms of elevation by season (?) -->

<!-- ## 15. Fix rate again (why?) -->
<!-- ```{r} -->
<!-- # Now restrict to the mode10 individuals ---------------------------------- -->
<!-- load("data/toKeep_fixrate.Rda") -->
<!-- toKeep_fixrate <- toKeep_fixrate[-1] -->
<!-- seasons_mode10 <- map2(seasons, toKeep_fixrate, ~.x %>% filter(Nili_id %in% .y)) -->

<!-- #save(seasons, file = "data/seasons.Rda") -->
<!-- #save(seasons_mode10, file = "data/seasons_mode10.Rda") -->
<!-- load("data/seasons.Rda") -->
<!-- load("data/seasons_mode10.Rda") -->
<!-- ``` -->
<!-- Products: seasons, seasons_mode10 -->
<!-- Metrics: Number of individuals per season (I don't think this should have changed from the step before? I'm not clear on why we are doing this again) -->

<!-- ## 16. Downsample to 10 minutes to deal with the more frequent fixes -->
<!-- ```{r} -->
<!-- # Downsample the data and save the downsampled data -->
<!-- subsample <- function(df, idCol = "Nili_id", timestampCol = "timestamp", mins = 10){ -->
<!--   sub <- df %>% -->
<!--     arrange(idCol, timestampCol) %>% -->
<!--     group_by(.data[[idCol]]) %>% -->
<!--     mutate(tc = cut(.data[[timestampCol]],  -->
<!--                     breaks = paste(as.character(mins), "min", sep = " "))) %>% -->
<!--     group_by(.data[[idCol]],  -->
<!--              "d" = lubridate::date(.data[[timestampCol]]), tc) %>% -->
<!--     slice(1) %>% -->
<!--     ungroup() %>% -->
<!--     dplyr::select(-c("d", "tc")) -->
<!--   return(sub) -->
<!-- } -->

<!-- # create and save downsampled datasets for later use -->
<!-- seasons_10min <- map(seasons, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) # basically just trying to remove any data from the occasional bursts of more concentrated and frequent sampling. -->
<!-- seasons_mode10_10min <- map(seasons_mode10, ~subsample(df = .x, idCol = "Nili_id", timestampCol = "timestamp", mins = 10), .progress = T) -->

<!-- map_dbl(seasons, nrow) -->
<!-- map_dbl(seasons_10min, nrow) -->
<!-- map_dbl(seasons_mode10, nrow) -->

<!-- seasons_mode10_10min <- map(seasons_mode10_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly)))) -->
<!-- seasons_10min <- map(seasons_10min, ~.x %>% group_by(Nili_id) %>% mutate(daysTracked = length(unique(dateOnly)))) -->

<!-- save(seasons_10min, file = "data/seasons_10min.Rda") -->
<!-- save(seasons_mode10_10min, file = "data/seasons_mode10_10min.Rda") -->
<!-- ``` -->