### 2024-02-28 temporal.qmd
more work with reducibility and temporal analysis in advance of meeting with Sandra

---
title: "Examining vulture social networks across time scales"
author: "Kaija Gahm"
date: 2024-02-28
format: 
  html:
    self-contained: true
execute:
  message: false
  warning: false
editor: visual
---

```{r}
#| include: false
library(tidyverse)
library(here)
theme_set(theme_classic())
```

```{r}
#| include: false
# Graphs of metrics -------------------------------------------------------
load(here::here("data/metrics_indiv.Rda"))
load(here::here("data/derived/cc.Rda"))
glimpse(metrics_indiv)
metrics_indiv <- metrics_indiv %>%
  group_by(ndays, int) %>%
  mutate(normDegree = degree/(n-1),
         normStrength = strength/sum(strength, na.rm = T)) # XXX WHAT IS WRONG HERE?
```

```{r}
#| echo: false
metrics_indiv %>%
  ggplot(aes(x = ndays, y = normDegree, col = type))+
  ylab("Degree (normalized)")+
  xlab("Time window (days)")+
  geom_jitter(alpha = 0.1, width = 0.6)+
  theme(axis.title = element_text(size = 14))+
  facet_wrap(~type) # There are a few individuals that don't roost with the others and therefore have a low degree, even at the high time windows. I thought those were Carmel birds, but I've now removed the northern indivs and we still see this pattern.

metrics_indiv %>%
  ggplot(aes(x = ndays, y = normStrength, col = type))+
  geom_jitter(alpha = 0.1, width = 0.6)+
  xlab("Time window (days)")+
  ylab("Strength (normalized)")+
  theme(axis.title = element_text(size = 14))+
  facet_wrap(~type) # I do not understand why this goes DOWN. Am I normalizing it incorrectly?

# How do network-level measures change over the time points?
metrics_net <- metrics_indiv %>%
  group_by(ndays, int, type, n) %>%
  summarize(mndeg = mean(degree),
            mnstr = mean(strength),
            mnnormdeg = mean(normDegree),
            mnnormstr = mean(normStrength)) %>%
  mutate(int = lubridate::ymd(int))
```

```{r}
#| echo: false
metrics_net %>%
  ggplot(aes(x = int, y = mnnormdeg, col = factor(ndays)))+
  ylab("Mean degree (normalized)")+
  xlab("Date")+
  geom_point()+
  geom_line()+
  facet_wrap(~type, ncol = 1, scales = "free")+
  theme(axis.title = element_text(size = 14))+
  labs(color = "Time window\n(days)")+
  geom_hline(aes(yintercept = 0.5), col = "black", linetype = 2)
```

```{r}
#| echo: false
#| fig-height: 10
metrics_net %>%
  ggplot(aes(x = mnnormdeg, col = type))+
  xlab("Mean degree (normalized)")+
  geom_density(linewidth = 1)+
  facet_wrap(~interaction(type, ndays), scales = "free", ncol = 3, nrow = 7)+
  scale_color_manual(name = "Situation",
                     values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14))+
  geom_vline(aes(xintercept = 0.5), linetype = 2)
```

```{r}
#| echo: false
metrics_net %>%
  group_by(ndays, type) %>%
  summarize(mn_mnnormdeg = mean(mnnormdeg)) %>%
  ggplot(aes(x = ndays, y = mn_mnnormdeg, col = type))+
  geom_point(size = 5)+
  geom_line(linewidth = 2)+
  ylab("Mean degree (normalized)")+
  xlab("Time window (days)")+
  scale_color_manual(name = "Situation",
                     values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  geom_hline(aes(yintercept = 0.5), linetype = 2)+
  theme(text = element_text(size = 14))
```