### 2024-03-29 scripts/00.1_wrapAround.qmd
brought over wrap-around code

---
title: "Wrap Around"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
```

## Load data

```{r}
load
```

```{r}
sms <- seq(from = 1, to = 25, by = 1)

sms_sims_conveyor <- vector(mode = "list", length = length(sms))
for(i in 1:length(sms)){
  shift <- sms[i]
  cat("shiftmax = ", shift, "\n")
  sims_conveyor <- map(sims_xy, ~{
    realizations <- vector(mode = "list", length = n)
    for(j in 1:n){
      cat(".")
      realizations[[j]] <- rotate_data_table(dataset = .x, shiftMax = shift,
                                             idCol = "indiv", dateCol = "date",
                                             timeCol = "time")
    }
    cat("\n")
    return(realizations)
  })
  sms_sims_conveyor[[i]] <- sims_conveyor
}
save(sms_sims_conveyor, file = "data/simulations/sms_sims_conveyor.Rda")
load("data/simulations/sms_sims_conveyor.Rda")
```

## Create wrapped data