### 2024-04-01 scripts/00.1_wrapAround.qmd
wrapping around and calculating graph metrics

---
title: "Wrap Around"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(future)
source(here("scripts/00.0_functions.R"))
roostPolygons <- sf::st_read(here("data/raw/roosts50_kde95_cutOffRegion.kml"))
```

## Load data

```{r}
load(here("data/dataPrep/downsampled_10min_forSocial.Rda"))
```

It's completely not computationally tractable to first make the entire dataset with 100 replicates and then calculate networks on it. Instead, we're going to have to iterate through, do each one, calculate the networks on each one, calculate the metrics, and save those.

```{r}
# for each season (map2 data, season_names)
  # make list for metrics for each rep
    # for each rep"
      # rotate
      # calc sri on the rotated data
      # make graph from the sri
      # calculate graph metrics
      # save graph metrics to spot in list
      # delete rotated data, sri, graph, and metrics
  # Bind the metrics with the season name .y

test <- downsampled_10min_forSocial[[1]] %>% 
  mutate(time = substr(as.character(timestamp), 12, 19))
roosts <- vultureUtils::get_roosts_df(test, id = "Nili_id") %>%
  sf::st_as_sf(., crs = "WGS84", coords = c("location_long", "location_lat"), remove = F)
reps <- 100
shiftMax <- 2
fn <- function(test, iter){
  # rotate data
  message("rotating")
  rotated <- rotate_data_table(dataset = test, shiftMax = shiftMax, 
                                   idCol = "Nili_id", dateCol = "dateOnly",
                                   timeCol = "time") %>% 
    mutate(rep = iter,
           timestamp = newdatetime) %>%
    sf::st_as_sf(., coords = c("location_long", "location_lat"), remove = F, crs = "WGS84")

  tojoin_roosts <- rotated %>% st_drop_geometry() %>% 
    select(Nili_id, "roost_date" = dateOnly, newDate) %>% distinct()
  roosts_temp <- roosts %>%
    left_join(tojoin_roosts, by = c("Nili_id", "roost_date")) %>%
    mutate(roost_date = newDate)

  # calculate sri for rotated data
  message("calculating sri")
  fl <- vultureUtils::getFlightEdges(rotated, roostPolygons = roostPolygons,
                        distThreshold = 1000, idCol = "Nili_id",
                        return = "sri")
  fe <- vultureUtils::getFeedingEdges(rotated, roostPolygons = roostPolygons,
                        distThreshold = 50, idCol = "Nili_id",
                        return = "sri")
  ro <- vultureUtils::getRoostEdges(roosts_temp, mode = "distance",
                                    distThreshold = 500, latCol = "location_lat",
                                    longCol = "location_long", idCol = "Nili_id",
                                    dateCol = "roost_date", return = "sri")

  # make graph from sri
  message("making graph")
  g_fl <- makeGraph(mode = "sri", data = fl, weighted = T)
  g_fe <- makeGraph(mode = "sri", data = fe, weighted = T)
  g_ro <- makeGraph(mode = "sri", data = ro, weighted = T)

  # calculate graph metrics
  message("calculating metrics")
  graph_metrics <- data.frame(degree = degree(g_fl),
                   strength = strength(g_fl),
                   Nili_id = names(degree(g_fl)),
                   situ = "flight") %>%
    bind_rows(data.frame(degree = degree(g_fe),
                         strength = strength(g_fe),
                         Nili_id = names(degree(g_fe)),
                         situ = "feeding")) %>%
    bind_rows(data.frame(degree = degree(g_ro),
                         strength = strength(g_ro),
                         Nili_id = names(degree(g_ro)),
                         situ = "roosting"))
  row.names(graph_metrics) <- NULL

  # save graph metrics to spot in list
  return(graph_metrics)
}

metrics <- map(1:reps, ~fn(test, iter = .x), .progress = T)
```