### 2024-04-15 presentation.qmd
moved .qmd docs to root

---
title: "Movement/Social"
date: last-modified
format:
  revealjs:
    embed-resources: true
editor: visual
---

```{r}
#| include: false
#| output: false
library(tidyverse)
library(here)
library(targets)
theme_set(theme_minimal())
tar_config_set(store = here::here('_targets'))
source(here("scripts/04.0_mixedModels.R"))
```

<!-- ## Research questions (cont'd) -->

<!-- Q1: What is the relationship between spatial phenotype and social interactions? -->

<!-- -   H1A: Individuals that move more will have more, and stronger, encounters with other unique individuals. -->

<!--     -   P1A: Vultures that move more will encounter more others (higher degree centrality in the social network) and have more interactions (higher strength in the social network) than vultures who move less. -->

<!-- ## Research questions (cont'd) -->

<!-- -   H1B: Individuals that are more exploratory will have more but weaker encounters with other unique individuals. -->

<!--     -   P1B: Exploratory vultures will encounter more unique individuals (higher degree centrality), but will have fewer interactions (lower strength) compared to less exploratory vultures. -->

<!-- ## Research questions (cont'd) -->

<!-- Q2: How is the relationship between spatial phenotype and social interactions impacted by social situation and season? -->

<!-- -   H2A: Social situations in which animals move more will have a greater impact on social interactions -->

<!--     -   P2A: Social centrality in co-flight networks will be more strongly associated with spatial phenotype than in co-roosting and co-feeding behavioral situations. -->

<!-- ## Research questions (cont'd) -->

<!-- -   H2B: Seasons in which animals move more will have a stronger impact on social interactions -->

<!--     -   P2B: The relationship between movement and social position will be stronger in the non-breeding seasons (summer and fall) than during the breeding season. -->

## Examining response variable distributions

```{r}
tar_load(all_movement_vars)
tar_load(cc)
all_movement_vars %>%
  pivot_longer(cols = c(movement, space_use, roost_div), 
               names_to = "var", values_to = "measure") %>%
  mutate(season = str_extract(seasonUnique, pattern = "[a-z]+")) %>%
  ggplot(aes(x = measure, col = season, group = seasonUnique))+
  geom_density()+
  facet_wrap(~var)+
  scale_color_manual(values = c(cc$breedingColor, cc$fallColor, cc$summerColor))+
  theme(legend.position = "bottom",
        axis.title = element_blank())
```

## Wrap-around randomizations (degree)

```{r}
#| fig-cap: "100 randomizations, shift +/- 1 day"
tar_load(allMetrics)
set.seed(3)
vultures <- sample(unique(allMetrics$Nili_id), 3, replace = F)
allMetrics %>%
  filter(Nili_id %in% vultures) %>%
  ggplot(aes(x = Nili_id, y = wrapped_normDegree, col = situ))+
  geom_point(aes(y = normDegree), position = position_dodge(width = 0.4))+
  geom_violin(position = position_dodge(width = 0.4))+
  facet_wrap(~season, scales = "free")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  ylab("Degree (normalized)")+
  theme(axis.title.x = element_blank(),
        legend.position = "bottom")+
  ggtitle("Degree (observed vs. random)")
```

## Wrap-around randomizations (strength)

```{r}
#| fig-cap: "100 randomizations, shift +/- 1 day"
allMetrics %>%
  filter(Nili_id %in% vultures) %>%
  ggplot(aes(x = Nili_id, y = wrapped_normStrength, col = situ))+
  geom_point(aes(y = normStrength), position = position_dodge(width = 0.4))+
  geom_violin(position = position_dodge(width = 0.4))+
  facet_wrap(~season, scales = "free")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  ylab("Strength (normalized)")+
  theme(axis.title.x = element_blank(),
        legend.position = "bottom")+
  ggtitle("Strength (observed vs. random)")
```

## Degree, but not strength, was higher than random.

```{r}
tar_load(linked)

linked %>%
  select(seasonUnique, situ, z_deg, z_str) %>%
  ungroup() %>%
  pivot_longer(cols = contains("z_"), names_to = "measure", values_to = "z") %>%
  ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
  geom_density()+
  facet_wrap(~measure, scales = "free")+
  geom_vline(aes(xintercept = 0), col = "red")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  theme(axis.title.y = element_blank())+
  xlab("Z-score")

linked %>%
  select(seasonUnique, situ, z_str) %>%
  ungroup() %>%
  ggplot(aes(x = z_str))
```

## Movement vs. degree deviation

```{r}
#| label: fig-d-space
#| fig-cap: "Relationship between movement and degree deviation"
plot_d_eff_movement
```

## Movement measure

```{r}
tar_load(movement_corr)
movement_corr
```

## Space use vs. degree deviation

```{r}
#| label: fig-d-space-situ
#| fig-cap: "Relationship between normalized degree and space use, with separate trend lines for each social situation"
plot_d_eff_situ_space
```

## Space use vs. degree deviation

```{r}
plot_d_emt_situ_space
```

## Space use measure

```{r}
tar_load(space_corr)
space_corr
```

## Roost diversification vs. degree deviation

```{r}
#| label: fig-d-roost
#| fig-cap: "Relationship between normalized degree and roost diversification"
plot_d_eff_roost 
```

## Roost diversification measure

```{r}
tar_load(roost_corr)
roost_corr
```

## Degree model summary {.smaller .scrollable}

```{r}
summary(degree_mod)
```

## The strength model did not fit well.

```{r}
simulationOutput <- DHARMa::simulateResiduals(strength_mod)
plot(simulationOutput, pch=".")
```

## Still, here is the summary as it stands {.smaller .scrollable}

```{r}
summary(strength_mod)
```