### 2024-04-17 presentation.qmd
updated presentation before and after 4/17 vulture meeting

---
title: "Movement/Social"
date: last-modified
date-format: long
format:
  revealjs:
    embed-resources: true
editor: visual
---

```{r}
#| include: false
#| output: false
library(tidyverse)
library(here)
library(targets)
library(patchwork)
library(ggeffects)
library(gtsummary)
theme_set(theme_classic())
tar_config_set(store = here::here('_targets'))
#source(here("scripts/04.0_mixedModels.R"))
tar_load(cc)
tar_load(linked)
tar_load(deg_mod)
tar_load(deg_z_mod)
tar_load(str_mod)
tar_load(str_z_mod)
tar_load(situcolors)
tar_load(seasoncolors)
tar_load(movement_corr)
tar_load(space_corr)
tar_load(roost_corr)
```

```{r}
#| include: false
effplot <- function(model, terms, dataset, pointX, pointY, colvar, legendTitle, values, ylab, xlab){
  dat <- as.data.frame(ggeffect(model, terms = terms))

  if(length(terms) > 1){
    plt <- ggplot(dat, aes(x, predicted)) +
      geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, linewidth = 0.6, show.legend = F)+
      geom_point(data = dataset %>% mutate("group" = .data[[colvar]]), 
                 aes(.data[[pointX]], .data[[pointY]], col = group), alpha = 0.5)+
      geom_line(aes(col = group), linewidth = 1) +
      facet_wrap(~group, scales = "free_y")+
      scale_color_manual(legendTitle, values = values) + 
      scale_fill_manual(legendTitle, values = values) + 
      ylab(ylab) + xlab(xlab)
  }else{
    plt <- ggplot(dat, aes(x, predicted)) +
      geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, linewidth = 0.6, show.legend = F)+
      geom_point(data = dataset %>% mutate("group" = .data[[colvar]]), 
                 aes(.data[[pointX]], .data[[pointY]], col = group), alpha = 0.5)+
      geom_line(linewidth = 1) +
      facet_wrap(~group, scales = "free_y")+
      scale_color_manual(legendTitle, values = values) + 
      ylab(ylab) + xlab(xlab)
  }
  return(plt)
}

emtplot <- function(model, specs, var, values, ylab, xlab){
  dat <- as.data.frame(emmeans::emtrends(model, specs, var))
  lower <- ifelse("lower.CL" %in% names(dat), "lower.CL", "asymp.LCL")
  upper <- ifelse("upper.CL" %in% names(dat), "upper.CL", "asymp.UCL")
  plt <- dat %>% ggplot(aes(.data[[specs]], 
                            .data[[paste0(var, ".", "trend")]], 
                            col = .data[[specs]])) + 
    geom_point(size = 6) +
    geom_errorbar(aes(x = .data[[specs]], 
                      ymin = .data[[lower]], 
                      ymax = .data[[upper]]), 
                  width = 0, linewidth = 2)+geom_hline(aes(yintercept = 0), 
                                                       linetype = 2)+
    scale_color_manual(values = values)+ 
    theme(legend.position = "none") + ylab(ylab)+ xlab(xlab) + coord_flip()
  return(plt)
}
```

## Movement

```{r}
movement_corr
```

## Space use

```{r}
space_corr
```

## Roost diversification

```{r}
roost_corr
```

## Predictor variable distributions

```{r}
linked %>%
  pivot_longer(cols = c(movement, space_use, roost_div), 
               names_to = "var", values_to = "measure") %>%
  mutate(season = str_extract(seasonUnique, pattern = "[a-z]+")) %>%
  ggplot(aes(x = measure, col = season, group = seasonUnique))+
  geom_density()+
  facet_wrap(~var)+
  scale_color_manual(values = c(cc$breedingColor, cc$fallColor, cc$summerColor))+
  theme(legend.position = "bottom",
        axis.title = element_blank())
```

# Predictor variable correlations

## Roost \~ Movement

```{r}
linked %>%
  select(movement, roost_div, space_use, season, seasonUnique, Nili_id) %>%
  distinct() %>%
  ggplot(aes(x = movement, y = roost_div, col = season, group = seasonUnique))+
  geom_point()+
  geom_smooth(method = "lm", alpha = 0.15)+
  scale_color_manual("Season", values = seasoncolors)+
  ylab("Roost diversification")+ xlab("Movement")
```

## Movement \~ Space

```{r}
linked %>%
  select(movement, roost_div, space_use, season, seasonUnique, Nili_id) %>%
  distinct() %>%
  ggplot(aes(x = space_use, y = movement, col = season, group = seasonUnique))+
  geom_point()+
  geom_smooth(method = "lm", alpha = 0.15)+
  scale_color_manual("Season", values = seasoncolors)+
  ylab("Movement")+ xlab("Space use")
```

## Space \~ Roost

```{r}
linked %>%
  select(movement, roost_div, space_use, season, seasonUnique, Nili_id) %>%
  distinct() %>%
  ggplot(aes(x = space_use, y = roost_div, col = season, group = seasonUnique))+
  geom_point()+
  geom_smooth(method = "lm", alpha = 0.15)+
  scale_color_manual("Season", values = seasoncolors)+
  ylab("Roost diversification")+ xlab("Space use")
```

## Response variable distributions (normalized degree)

```{r}
linked %>%
  ggplot(aes(x = normDegree, col = situ))+
  geom_density()+
  facet_wrap(~seasonUnique, scales = "free_y")+
  scale_color_manual("Situation", values = situcolors)+ylab("")
```

## Response variable distributions (normalized strength)

```{r}
linked %>%
  ggplot(aes(x = normStrength, col = situ))+
  geom_density()+
  facet_wrap(~seasonUnique, scales = "free_y")+
  scale_color_manual("Situation", values = situcolors)+ylab("")
```

## Wrap-around randomizations (degree)

```{r}
#| fig-cap: "100 randomizations, shift +/- 1 day"
tar_load(allMetrics)
set.seed(3)
vultures <- sample(unique(allMetrics$Nili_id), 3, replace = F)
allMetrics %>%
  filter(Nili_id %in% vultures) %>%
  ggplot(aes(x = Nili_id, y = wrapped_degree, col = situ))+
  geom_point(aes(y = degree), position = position_dodge(width = 0.4))+
  geom_violin(position = position_dodge(width = 0.4))+
  facet_wrap(~season, scales = "free")+
  scale_color_manual(name = "Situation", values = situcolors)+
  theme_classic()+
  ylab("Degree")+
  theme(axis.title.x = element_blank())+
  ggtitle("Degree (observed vs. random)")
```

## Wrap-around randomizations (strength)

```{r}
#| fig-cap: "100 randomizations, shift +/- 1 day"
allMetrics %>%
  filter(Nili_id %in% vultures) %>%
  ggplot(aes(x = Nili_id, y = wrapped_strength, col = situ))+
  geom_point(aes(y = strength), position = position_dodge(width = 0.4))+
  geom_violin(position = position_dodge(width = 0.4))+
  facet_wrap(~season, scales = "free")+
  scale_color_manual(name = "Situation", values = situcolors)+
  theme_classic()+
  ylab("Strength")+
  theme(axis.title.x = element_blank())+
  ggtitle("Strength (observed vs. random)")
```

## Response variable distributions (z-scores)

```{r}
linked %>%
  select(seasonUnique, situ, z_deg, z_str) %>%
  ungroup() %>%
  pivot_longer(cols = contains("z_"), names_to = "measure", values_to = "z") %>%
  ggplot(aes(x = z, col = situ, group = interaction(seasonUnique, measure, situ)))+
  geom_density(aes(linetype = measure))+
  facet_wrap(~measure*situ, scales = "free")+
  geom_vline(aes(xintercept = 0), col = "red")+
  scale_color_manual(name = "Situation", values = c(cc$feedingColor, cc$flightColor, cc$roostingColor))+
  theme_classic()+
  theme(axis.title.y = element_blank())+
  xlab("Z-score")
```

# Model results: Degree (z-scores, non-normalized)

## Space use and situation

```{r}
#| fig-cap: "Degree z-scores"
deg_z_situ_space_emt_P <- effplot(model = deg_z_mod, terms = c("space_use", "situ"), 
                            dataset = linked, pointX = "space_use", 
                            pointY = "z_deg", col = "situ", 
                            legendTitle = "Situation", values = situcolors, 
                            ylab = "Degree (z-score)", xlab = "Space use (log)")

deg_z_situ_space_P <- emtplot(model = deg_z_mod, specs = "situ", 
                                var = "space_use", values = situcolors, 
                                ylab = "Space use effect (Degree z-scores)", 
                                xlab = "Situation")

deg_z_situ_space_emt_P + deg_z_situ_space_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
```

##

![](fig/whiteboards/deg_z_space_situ_ro.jpg)

##

![](fig/whiteboards/deg_z_space_situ_fefl.jpg)

## Roost diversification

```{r}
#| fig-cap: "Degree z-scores"
deg_z_roost_P <- effplot(model = deg_z_mod, terms = c("roost_div", "situ"), 
                       dataset = linked, pointX = "roost_div", 
                       pointY = "z_deg", col = "situ", 
                       legendTitle = "Situation", values = situcolors, 
                       ylab = "Degree (z-score)", xlab = "Roost diversification")
deg_z_roost_P & theme_classic()
```

##

![](fig/whiteboards/deg_z_roost.jpg)

# Model results: Strength (z-scores, non-normalized)

## Space use and situation

```{r}
#| fig-cap: "Strength z-scores"
str_z_situ_space_emt_P <- emtplot(model = str_z_mod, specs = "situ", 
                                var = "space_use", values = situcolors, 
                                ylab = "Space use effect (strength z-score)", 
                                xlab = "Situation")

str_z_situ_space_P <- effplot(model = str_z_mod, terms = c("space_use", "situ"), 
                            dataset = linked, pointX = "space_use", 
                            pointY = "z_str", col = "situ", 
                            legendTitle = "Situation", values = situcolors, 
                            ylab = "Strength (z-score)", xlab = "Space use (log)")

str_z_situ_space_P + str_z_situ_space_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
```

## Roost diversification

```{r}
#| fig-cap: "Strength z-scores"
str_z_roost_P <- effplot(model = str_z_mod, terms = c("roost_div", "situ"), 
                         dataset = linked, pointX = "roost_div", 
                         pointY = "z_str", col = "situ", 
                         legendTitle = "Situation", values = situcolors, 
                         ylab = "Strength (z-score)", xlab = "Roost diversification")
str_z_roost_P & theme_classic()
```

# Model results: Normalized degree

## Movement and situation

```{r}
#| fig-cap: "Normalized degree"
deg_situ_movement_P <- effplot(model = deg_mod, terms = c("movement", "situ"), 
                               dataset = linked, pointX = "movement", 
                               pointY = "normDegree", col = "situ", 
                               legendTitle = "Situation", values = situcolors, 
                               ylab = "normDegree", xlab = "Movement")

deg_situ_movement_emt_P <- emtplot(model = deg_mod, specs = "situ", 
                                   var = "movement", values = situcolors, 
                                   ylab = "Movement effect (normDegree)", 
                                   xlab = "Situation")

deg_situ_movement_P + deg_situ_movement_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
```

## Movement and season

```{r}
#| fig-cap: "Normalized degree"
deg_season_movement_P <- effplot(model = deg_mod, terms = c("movement", "season"), 
                                 dataset = linked, pointX = "movement", 
                                 pointY = "normDegree", col = "season", 
                                 legendTitle = "Season", values = seasoncolors, 
                                 ylab = "normDegree", xlab = "Movement")

deg_season_movement_emt_P <- emtplot(model = deg_mod, specs = "season", 
                                     var = "movement", values = seasoncolors, 
                                     ylab = "Movement effect (normDegree)", 
                                     xlab = "Season")

deg_season_movement_P + deg_season_movement_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
```

## Space use and situation

```{r}
#| fig-cap: "Normalized degree"

deg_situ_space_P <- effplot(model = deg_mod, terms = c("space_use", "situ"), 
                            dataset = linked, pointX = "space_use", 
                            pointY = "normDegree", col = "situ", 
                            legendTitle = "Situation", values = situcolors, 
                            ylab = "normDegree", xlab = "Space use (log)")

deg_situ_space_emt_P <- emtplot(model = deg_mod, specs = "situ", 
                                var = "space_use", values = situcolors, 
                                ylab = "Space use effect (normDegree)", 
                                xlab = "Situation")

deg_situ_space_P + deg_situ_space_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
```

## Roost diversification

```{r}
#| fig-cap: "Normalized degree"
deg_roost_P <- effplot(model = deg_mod, terms = c("roost_div", "situ"), 
                               dataset = linked, pointX = "roost_div", 
                               pointY = "normDegree", col = "situ", 
                               legendTitle = "Situation", values = situcolors, 
                               ylab = "normDegree", xlab = "Roost diversification")
deg_roost_P & theme_classic()
```

# Model results: Normalized strength

## Roost diversification and situation

```{r}
#| fig-cap: "Normalized strength"
str_situ_roost_P <- effplot(model = str_mod, terms = c("roost_div", "situ"), 
                               dataset = linked, pointX = "roost_div", 
                               pointY = "normStrength", col = "situ", 
                               legendTitle = "Situation", values = situcolors, 
                               ylab = "normStrength", xlab = "Roost diversification")

str_situ_roost_emt_P <- emtplot(model = str_mod, specs = "situ", 
                                   var = "roost_div", values = situcolors, 
                                   ylab = "Roost diversification effect (normStrength)", 
                                   xlab = "Situation")
str_situ_roost_P + str_situ_roost_emt_P + plot_layout(widths = c(12, 4), guides = "collect") & theme_classic()
```

## Takeaways

1.  Movement relates to strength/degree, but much of this is incidental--effect drops away after randomizing
2.  Differences in excess/preferential sociality relate mostly to space use

# Model summaries

## Normalized (not z-scores) {.scrollable}

```{r}
tbl_deg_mod <- tbl_regression(deg_mod)
tbl_str_mod <- tbl_regression(str_mod)

merged <- tbl_merge(
  tbls <- list(tbl_deg_mod, tbl_str_mod),
  tab_spanner = c("**Degree (normalized)**", "**Strength (normalized)**")
)
merged

# save the table (requires the `gt` package)
#gt::gtsave(as_gt(merged), file = file.path("fig", "mm_summary.png"))
```

## Z-scores {.scrollable}

```{r}
tbl_deg_z_mod <- tbl_regression(deg_z_mod)
tbl_str_z_mod <- tbl_regression(str_z_mod)

merged_z <- tbl_merge(
  tbls <- list(tbl_deg_z_mod, tbl_str_z_mod),
  tab_spanner = c("**Degree (z-score)**", "**Strength (z-score)**")
)
merged_z

# save the table (requires the `gt` package)
#gt::gtsave(as_gt(merged_z), file = file.path("fig", "mm_summary_z.png"))
```